---
title: "Cleaning Data"
output: html_document
author: Jordan Koster
---
In this notebook, I import and clean the three datasets I'll be using for this project. 

Loading needed packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fs)
library(skimr)
library(janitor)
library(lubridate)
```
First I'll import the first dataset - EPA air pollution data from 2020
```{r}
EPA_2020 <- read_csv(path("..", "raw data", "EPA AQI by county 2020.csv"))
```
I want to get an idea of what the dataset looks like
```{r}
glimpse(EPA_2020)
skim(EPA_2020)
view(EPA_2020)
```
I'll clean up the names of the columns in the dataset to make them more user-friendly (all lowercase and joined words) 
```{r}
EPA_2020 <- clean_names(EPA_2020)
view(EPA_2020)
```
Filter the dataset so that we have only New York State data 
```{r}
EPA_2020 <- filter(EPA_2020, state_name == "New York")
view(EPA_2020)
```
Load in the second dataset (EPA air pollution data from 2021)
```{r}
EPA_2021 <- read_csv(path("..", "raw data", "EPA AQI by county 2021.csv"))
```
Get a look at the second dataset
```{r}
glimpse(EPA_2021)
skim(EPA_2021)
view(EPA_2021)
```
Clean up this dataset's column names too 
```{r}
EPA_2021 <- clean_names(EPA_2021)
view(EPA_2021)
```
Filter this dataset so we have only NYS data
```{r}
EPA_2021 <- filter(EPA_2021, state_name == "New York")
view(EPA_2021)
```
Get important info about both NYS-only datasets 
```{r}
skim(EPA_2020)
skim(EPA_2021)
```
Load in the final dataset I'll be using (NYS COVID fatalities data)
```{r}
NYS_Covid_Deaths <- read_csv(path("..", "raw data", "NYS COVID deaths by county.csv"))
```
Get a look at this dataset
```{r}
glimpse(NYS_Covid_Deaths)
view(NYS_Covid_Deaths)
```
Clean up this dataset's column names
```{r}
NYS_Covid_Deaths <- clean_names(NYS_Covid_Deaths)
view(NYS_Covid_Deaths)
```
Get important info about this dataset
```{r}
skim(NYS_Covid_Deaths)
```
Now I'll get rid of columns I won't need in these datasets 
```{r}
EPA_2020 <- select(EPA_2020, -state_name, -state_code, -defining_site)
EPA_2020
EPA_2021 <- select(EPA_2021, -state_name, -state_code, -defining_site)
EPA_2021
NYS_Covid_Deaths <- select(NYS_Covid_Deaths, -place_of_fatality)
NYS_Covid_Deaths
```
getting rid of a few more columns I won't need
```{r}
EPA_2020 <- select(EPA_2020, -county_code, -category)
EPA_2021 <- select(EPA_2021, -county_code, -category)
```

```{r}
view(EPA_2020)
view(EPA_2021)
view(NYS_Covid_Deaths)
```

This project will be using data for only Ozone and PM2.5 pollutants so now I'll get rid of the data for other pollutants in the 2 EPA datasets 
```{r}
EPA_2020 <- filter(EPA_2020, defining_parameter == "Ozone" | defining_parameter == "PM2.5")
view(EPA_2020)
EPA_2021 <- filter(EPA_2021, defining_parameter == "Ozone" | defining_parameter == "PM2.5")
view(EPA_2021)
```

Now I need to rename some of the columns so that they are common across the three datasets and easier to use
```{r}
EPA_2020 <- rename(EPA_2020, pollutant = defining_parameter)
EPA_2020 <- rename(EPA_2020, county = county_name)
view(EPA_2020)
EPA_2021 <- rename(EPA_2021, pollutant = defining_parameter)
EPA_2021 <- rename(EPA_2021, county = county_name)
view(EPA_2021)
NYS_Covid_Deaths <- rename(NYS_Covid_Deaths, date = report_date)
NYS_Covid_Deaths <- rename(NYS_Covid_Deaths, deaths = deaths_by_county_of_residence)
view(NYS_Covid_Deaths)
```
Making sure that the date columns in each dataset are in date format 
```{r}
glimpse(EPA_2020)
glimpse(EPA_2021)
glimpse(NYS_Covid_Deaths)
```
The date columns are not in the correct format so I need to reformat them 
```{r}
new_EPA_2021 <- EPA_2021 %>%
  mutate(new_date = mdy(date))
view(new_EPA_2021)
```
```{r}
new_NYS_Covid_Deaths <- NYS_Covid_Deaths %>%
  mutate(new_date = mdy(date))
view(new_NYS_Covid_Deaths)
new_EPA_2020 <- EPA_2020 %>%
  mutate(new_date = mdy(date))
view(new_EPA_2020)
```
correcting column names
```{r}
new_EPA_2021 <- select(new_EPA_2021, -date)
new_EPA_2021 <- rename(new_EPA_2021, date = new_date)
view(new_EPA_2021)
```
```{r}
new_NYS_Covid_Deaths <- select(new_NYS_Covid_Deaths, -date)
new_NYS_Covid_Deaths <- rename(new_NYS_Covid_Deaths, date = new_date)
view(new_NYS_Covid_Deaths)
new_EPA_2020 <- select(new_EPA_2020, -date)
new_EPA_2020 <- rename(new_EPA_2020, date = new_date)
view(new_EPA_2020)
```
Now all 3 datasets' date columns are in the correct format 

All 3 datasets are now clean 

the clean datasets' names are: 
new_EPA_2020 
new_EPA_2021
new_NYS_Covid_Deaths

#add code that automatically saves cleaned datasets to clean data folder


