---
title: "Merge Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Loading in needed packages 
```{r}
library(tidyverse)
library(fs)
library(skimr)
library(janitor)
```
Now that the three datasets are clean, I need to merge them into one dataset so that I can begin my analysis 

First I can combine the new_EPA_2020 and the new_EPA_2021 datasets. I'm using the rbind function because these two datasets have the same column names and are essentially just the 2020 and 2021 sets of the same collected data. 
```{r}
combined_data <- rbind(new_EPA_2020, new_EPA_2021)
view(combined_data)
```
Next I need to merge in the new_NYS_Covid_Deaths dataset. I'm using the inner_join() command because I want to look at only the time period that both the EPA data and the NYS Covid data  covers. This is the time period from 3/27/2020 to 3/31/2021. 
```{r}
merged_data <- inner_join(
  combined_data,
  new_NYS_Covid_Deaths,
  by = c("county", "date"),
  copy = FALSE,
  suffix = c(".x", ".y"),
  keep = FALSE,
  na_matches = c("na", "never")
)
view(merged_data)
```
Now I have one dataset (called "merged_data") that I can now use for my analysis going forward.
