---
title: "03_Regression Analysis and Visualization"
output: html_document
author: Jordan Koster
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
loading needed packages 
```{r}
library(tidyverse)
library(fs)
library(skimr)
library(ggpubr)
library(broom)
library(ggpmisc)
library(ggplot2)
library(gtsummary)
library(gt)
```
Now that my data is cleaned and has been merged into one dataset, I can start my analysis 

My main question coming into this project was: How have pollution levels changed over the course of the COVID-19 pandemic? Have the pandemic lockdown policies affected these pollution levels?
(pollution levels = AQI (Air Quality Index) measurements by EPA)

I'm interested in how lockdown policies affected the amount of air pollution, specifically in NYS. I'm using COVID deaths as a proxy for lockdown policies with the assumption that more deaths means stricter lockdown policies. My hypothesis is that as lockdown policies become more strict (proxy: as COVID deaths increase), pollution levels (AQI measurements) will decrease. 

So, to explore the relationship between the number of COVID deaths (the "deaths" column of the dataset) and the amount of air pollution (the "aqi" column of the dataset), I'm running a regression. 

First, I want to run a regression with deaths as my independent variable and aqi (air pollution) as my dependent variable without adding in any controls. 
```{r}
regression_results <- lm(aqi ~ deaths, data = merged_data) %>% 
    tidy()
view(regression_results)
```
I got a p value of 0.2435661. It is not statistically significant at the 5% level. 

Now, I'll run the same regression but also add in controls. Because I am working with a fairly simple dataset that does not have many variables, I am limited in how much I can control for. 

I'm going to control for county, date, pollutant, and number_of_sites_reporting

```{r}
regression_results_with_controls <- lm(aqi ~ deaths + county + date + pollutant + number_of_sites_reporting, data = merged_data) %>% 
    tidy()
view(regression_results_with_controls)
```
I got a p value of 2.767511e-23. It is significant at the 1% level.

With my hypothesis stated above, I was expecting a negative relationship between the two variables of interest, so my estimate should be negative. However, the estimate given by the regression is 0.003324804, indicating a positive relationship. 

To explore this positive relationship, I'm going to run a regression with deaths as my dependent variable and aqi (air pollution) as my independent variable with same controls as above. 
```{r}
reverse_regression_results_with_controls <- lm(deaths ~ aqi + county + date + pollutant + number_of_sites_reporting, data = merged_data) %>% 
    tidy()
view(reverse_regression_results_with_controls)
```
I got a p value of 2.767511e-23, which is significant at the 1% level. The estimate is 3.187643, indicating a positive relationship. 

This was not the result I was expecting but it makes sense (see my blog post for more information)

Visualizations  

Visualizing first regression (lines 44-46)
#Thank you to Keith Wescott for his code to make this visualization 
```{r}
my.formula <- y ~ x
ggplot(data = merged_data) +
  aes(x = deaths, y = aqi, color = pollutant) +
  geom_point(size = 0.5, alpha = 0.25) +
  geom_smooth(method=lm, se=TRUE, show.legend=TRUE,formula = my.formula) +
   stat_poly_eq(formula = my.formula, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) + 
  coord_cartesian(ylim = c(0,100)) +
  labs(
    title = "Regression of AQI on COVID Deaths",
    subtitle = "By Pollutant",
    x = "COVID Deaths",
    y = "Pollutant AQI levels",
    caption = "Source: EPA and NYS Health",
  )
```
Visualizing second (reverse) regression (lines 54-56 )
```{r}
my.formula.1 <- y ~ x
ggplot(data = merged_data) +
  aes(x = aqi, y = deaths, color = pollutant) +
  geom_point(size = 0.5, alpha = 0.25) +
  geom_smooth(method=lm, se=TRUE, show.legend=TRUE,formula = my.formula.1) +
   stat_poly_eq(formula = my.formula.1, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) + 
  coord_cartesian(ylim = c(0,4000), xlim = c(0,100)) +
  labs(
    title = "Regression of COVID Deaths on AQI",
    subtitle = "By Pollutant",
    x = "Pollutant AQI Levels",
    y = "COVID Deaths",
    caption = "Source: EPA and NYS Health",
  )
```
Table showing first regression results
```{r}
first_regression <- lm(
  aqi ~ deaths + county + date + pollutant + number_of_sites_reporting, 
  data = merged_data,
)
  
tbl_regression(
  first_regression,
  label = list(
    "deaths" ~ "COVID Deaths in NYS", 
    "county" ~ "County in NYS", 
    "date" ~ "Date",
    "pollutant" ~ "Ozone and PM2.5",
    "number_of_sites_reporting" ~ "Sites where AQI measured"
  )
)
```
Table showing second (reverse) regression results
```{r}
reverse_regression <- lm(
  deaths ~ aqi + county + date + pollutant + number_of_sites_reporting, 
  data = merged_data,
)
  
tbl_regression(
  reverse_regression,
  label = list(
    "aqi" ~ "AQI Pollutant Measurements", 
    "county" ~ "County in NYS", 
    "date" ~ "Date",
    "pollutant" ~ "Ozone and PM2.5",
    "number_of_sites_reporting" ~ "Sites where AQI measured"
  )
)
```


