---
title: "03_Regression Analysis and Visualization"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
loading needed packages 
```{r}
library(tidyverse)
library(fs)
library(skimr)
library(ggpubr)
library(broom)
library(ggpmisc)
library(ggplot2)
library(gtsummary)
library(gt)
```
Now that my data is cleaned and has been merged into one dataset, I can start my analysis 

My main question coming into this project was: How have pollution levels changed over the course of the COVID-19 pandemic? Have the pandemic lockdown policies affected these pollution levels?
(pollution levels = AQI (Air Quality Index) measurements by EPA)

I'm interested in how lockdown policies affected the amount of air pollution, specifically in NYS. I'm using COVID deaths as a proxy for lockdown policies with the assumption that more deaths means stricter lockdown policies. My hypothesis is that as lockdown policies become more strict (proxy: as COVID deaths increase), pollution levels (AQI measurements) will decrease. 

So, to explore the relationship between the number of COVID deaths (the "deaths" column of the dataset) and the amount of air pollution (the "aqi" column of the dataset), I'm running a regression. 

First, I want to run a regression with deaths as my independent variable and aqi (air pollution) as my dependent variable without adding in any controls. 
```{r}
regression_results <- lm(aqi ~ deaths, data = merged_data) %>% 
    tidy()
view(regression_results)
```
I got a p value of 0.2435661. It is not statistically significant at the 5% level. 

Now, I'll run the same regression but also add in controls. Because I am working with a fairly simple dataset that does not have many variables, I am limited in how much I can control for. 

I'm going to control for county, date, pollutant, and number_of_sites_reporting

```{r}
regression_results_with_controls <- lm(aqi ~ deaths + county + date + pollutant + number_of_sites_reporting, data = merged_data) %>% 
    tidy()
view(regression_results_with_controls)
```
I got a p value of 2.767511e-23. so significant but positive relationship (expecting negative relationship)


Visualization  

#Thank you to Keith Wescott for his code to make this visualization 
```{r}
my.formula <- y ~ x
ggplot(data = merged_data) +
  aes(x = deaths, y = aqi, color = pollutant) +
  geom_point(size = 0.5, alpha = 0.25) +
  geom_smooth(method=lm, se=TRUE, show.legend=TRUE,formula = my.formula) +
   stat_poly_eq(formula = my.formula, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
                parse = TRUE) + 
  coord_cartesian(ylim = c(0,156)) +
  labs(
    title = "Regression of AQI on COVID Deaths",
    subtitle = "By Pollutant",
    x = "COVID Deaths",
    y = "Pollutant AQI levels",
    caption = "Source: EPA and NYS Health",
  )
```
#cool table - still figuring it out 
```{r}
test_regression <- lm(
  aqi ~ deaths + county + date + pollutant + number_of_sites_reporting, 
  data = merged_data,
)
  
tbl_regression(
  test_regression,
  label = list(
    "deaths" ~ "COVID Deaths in NYS", 
    "county" ~ "County in NYS", 
    "date" ~ "Date",
    "pollutant" ~ "Ozone and PM2.5",
    "number_of_sites_reporting" ~ "Sites where AQI measured"
  )
)
   
```


